Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-11 paper/paper.tex
Updated organization of the paper

The implementation  of our  analysis annotates each  potential garbage
collection  point  in   the  program  with  a   set  of  deterministic
finite-state automata (DFA) describing the  liveness at that point. We
also implement a copying collector that  uses the DFA to preserve live
objects.   Our  experiments  confirm  that  liveness-based  collection
always increases  the garbage  reclaimed, consequently  decreasing the
number of collections.  In addition,  it reduces the execution time of
several programs.  Experiments  also reveal that the  best results are
obtained  for   a  mixed  garbage   collection  scheme  that   does  a
liveness-based    collection   for    fully    evaluated   data    and
reachability-based   collection  for   closures.   On   our  benchmark
programs,  we  obtain a  decrease  of  up to  45\%  in  the number  of
collections, and  a decrease of  up to 88\%  in the number  of dragged
cells.  Even  though  liveness   based  collector  uses  an  expensive
traversal scheme,  the execution  time is  competitive (0.2X  - 2.08X)
when compared to a reachability based collector.
\end{abstract}

@@

\subsection{Results}
\newcommand{\hgt}{3cm}
\begin{figure*}[p]
\renewcommand{\arraystretch}{.1}
\begin{tabular}{@{}c@{}@{}c@{}}
   \hskip -4mm{\epsfig{file=lambda.eps, height=\hgt}}
   & {\epsfig{file=lambda_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nperm.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nperm_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=treejoin.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=treejoin_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=lcss.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=lcss_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=sudoku.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=sudoku_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=fibheap.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=fibheap_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nqueens.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nqueens_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=knightstour.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=knightstour_win.eps, height=\hgt}}
\\ (a) & (b)
\end{tabular}%\vskip -10mm
 \caption{Memory usage.  Column (a) shows complete  usage while column
   (b) enlarges  a part  of the  usage.  The blue  and the  red curves
   indicate the number of cons cells  in the active semi-space for RGC
   and LGC  respectively.  The  black curve  represents the  number of
   reachable cells and  the grey curve represents the  number of cells
   that are  actually live (of  which liveness analysis does  a static
   approximation).   x-axis   is  the  time  measured   in  number  of
   cons-cells allocated (scaled down by  factor $10^5$). y-axis is the
   number of cons-cells (scaled down by factor $10^3$).}
\label{fig:memory-usage} \figrule
\end{figure*}

>>>>

The implementation  of our  analysis annotates each  potential garbage
collection  point   in  the  program  with  a   set  of  deterministic
finite-state automata (DFA) describing  the liveness at that point. We
also implement a copying collector  that uses the DFA to preserve live
objects.   Our  experiments  confirm  that  liveness-based  collection
always  increases the garbage  reclaimed, consequently  decreasing the
number of collections.  In addition,  it reduces the execution time of
several programs.  Experiments also reveal  that the best  results are
obtained  for   a  mixed  garbage   collection  scheme  that   does  a
liveness-based    collection   for    fully    evaluated   data    and
reachability-based   collection  for   closures.   On   our  benchmark
programs,  we  obtain a  decrease  of  up to  50\%  in  the number  of
collections, and up to 70\% decrease in overall program execution time
when compared to reachability based collectors.
\end{abstract}

@@

\subsection{Results}
\newcommand{\hgt}{3cm}
\begin{figure*}[p]
\renewcommand{\arraystretch}{.1}
\begin{tabular}{@{}c@{}@{}c@{}}
   \hskip -4mm{\epsfig{file=lambda.eps, height=\hgt}}
   & {\epsfig{file=lambda_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nperm.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nperm_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=treejoin.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=treejoin_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=lcss.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=lcss_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=sudoku.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=sudoku_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=fibheap.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=fibheap_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=nqueens.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=nqueens_win.eps, height=\hgt}}
\\ \hskip -4mm{\epsfig{file=knightstour.eps, height=\hgt}}
&  \hskip -4mm{\epsfig{file=knightstour_win.eps, height=\hgt}}
\\ (a) & (b)
\end{tabular}%\vskip -10mm
 \caption{Memory  usage  for (a)  full  benchmarks  and (b)  for  some
   interesting part  of the benchmarks.   The blue and the  red curves
   indicate the number of cons cells  in the active semi-space for RGC
   and LGC  respectively.  The  black curve  represents the  number of
   reachable cells and  the grey curve represents the  number of cells
   that are  actually live (of  which liveness analysis does  a static
   approximation).   x-axis   is  the  time  measured   in  number  of
   cons-cells allocated (scaled down by  factor $10^5$). y-axis is the
   number of cons-cells (scaled down by factor $10^3$).}
\label{fig:memory-usage} \figrule
\end{figure*}
